\pdfoutput=1
\documentclass[11pt]{article}

% ACL style
\usepackage{acl}
\usepackage{natbib}
\usepackage{times}
\usepackage{latexsym}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{xurl}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage[utf8]{inputenc}
\usepackage[T5,T1]{fontenc}
\usepackage[vietnamese,english]{babel}

\title{Hate Speech Detection and Highlighting\\in Vietnamese Social Media Comments}
\author{
  \textbf{Phat Nguyen Cong}$^{1,2}$, \textbf{An Nguyen Xuan}$^{1,2}$, \textbf{Binh Mai Thai}$^{1,2}$, \\ % Ngắt hàng tại đây
  \textbf{An Truong Hoang Thanh}$^{1,2}$, \textbf{Huong Nguyen Le Quynh}$^{1,2}$ \\
  \addlinespace[0.2cm] % Tạo khoảng cách nhỏ giữa tên và địa chỉ
  $^1$University of Information Technology, Ho Chi Minh City, Vietnam \\
  $^2$Vietnam National University, Ho Chi Minh City, Vietnam \\
  \texttt{\{23521143, 23520023, 23520158, 23520032, 21520255\}@gm.uit.edu.vn}
}

\begin{document}
\maketitle

\begin{abstract}
The rise of hate speech on Vietnamese social media has created a real need for systems that
can understand deep context rather than just a simple classification. In this work, we tackle
the challenge of target-specific and implicit hate speech by introducing the HARE framework.
By integrating Chain-of-Thought (CoT) reasoning into the Qwen2.5-3B model, our approach
moves beyond traditional ``black-box'' models. Instead of just providing a classification label, HARE generates semantic explanations to reveal
the hidden intent behind a text, using a two-stage training process on the ViTHSD dataset.
Our experiments show that the model achieves an F1-score of 60.26\%, outperforming
benchmarks like PhoBERT and Flan-T5, especially in the politics category with a notable
24.87\% jump. Finally, we provide a practical architecture for deploying this solution on
large-scale streaming platforms like Apache Kafka and Spark, integrating keyword highlighting
features, proving that our approach is ready for real-world use.
\end{abstract}

\section{Introduction}

\subsection{Overview of Social Media Challenges in Vietnam}
In recent years, social media has become a central part of public life in Vietnam. Platforms
such as Facebook, YouTube, and TikTok are now the main places where people discuss
everyday topics. However, this openness has also led to a rise in toxic language, creating a
serious challenge for keeping these online spaces safe.

One major difficulty in detecting this content comes from the special features of the
Vietnamese language. Vietnamese is monosyllabic and tonal, and meaning often depends on
cultural context. Today, users often use insulting slang like \foreignlanguage{vietnamese}{\textit{``ba que''}} or \foreignlanguage{vietnamese}{\textit{``bò đỏ''}} along with
teencode, sarcasm, and wordplay. These creative ways of speaking help harmful content
avoid basic keyword filters, making it hard for both automated systems and human
moderators to control.

Therefore, using modern technologies such as pre-trained language models like PhoBERT still
has some limitations:
\begin{itemize}[leftmargin=*]
\item Lack of Explainability: Deep learning models often operate as ``black boxes''
producing classification labels without providing specific reasoning, which makes it
difficult to build trust and transparency in content moderation.
\item Limitation in recognizing implicit hate: Most current models focus so much on
surface-level keywords that they only catch obvious swearing. They often fail to detect
``hidden attacks'' because these insults rely on subtle comparisons and cultural context
rather than explicit hate speech.
\item Limitation in recognizing target: Most previous research has focused almost entirely
on how toxic a comment is while overlooking the specific target of the attack.
Identifying whether the victim is an individual, an organization, or an ethnic group is
actually a vital piece of information. Knowing exactly who is being targeted is essential
for assessing the severity of the speech and determining the most appropriate way to
intervene.
\end{itemize}

To address these practical challenges, we focus on building an automated system that can
explain both ``why'' a piece of content is considered hateful and ``who'' is being targeted. This
objective serves as the foundation for our work with the ViTHSD dataset and the modeling
approaches that we present in this paper.

\subsection{Objectives of the report}
This report aims to address the above limitations by developing a comprehensive system
based on the ViTHSD dataset. The key contributions include:
\begin{enumerate}[leftmargin=*]
\item Systematizing the Targeted Hate Speech Problem: We redefine the hate speech
detection problem in Vietnamese from a target-oriented perspective, using the ViTHSD
dataset with a multi-layered label structure.
\item Applying the HARE Framework and CoT: Proposing a process to integrate the
reasoning capabilities of LLMs to generate explanations (rationales) and implied
statements, helping the model better understand the context.
\item Two-Stage Semantic Alignment Training Strategy: Introducing the two-stage
training method on the Qwen2.5-3B model, which enables the model to learn decision
boundaries from labeled data and reasoning from rationale data.
\item Experiments and Evaluation: Providing experimental results comparing with the
strongest current baselines (PhoBERT, Flan-T5) and a deep analysis of the method's
effectiveness on challenging labels such as Politics and Group Hate.
\item Introducing the practical deployment architecture: Introducing an AI-integrated
system for real-time data streaming using Kafka and Spark, with FastAPI for the
backend and React and Vite for the frontend, designed for large-scale social media
monitoring applications.
\end{enumerate}

\section{Literature Review and Theoretical Background}

\subsection{The Development of Vietnamese NLP and Pre-trained Models}
Natural Language Processing (NLP) for Vietnamese has undergone a major transformation
over the last decade. Before 2019, most research relied on hand-crafted features and static
models like Word2Vec \citep{mikolov2013efficient} or FastText \citep{bojanowski2017enriching}, but the arrival of the Transformer architecture \citep{vaswani2017attention} and
Google's BERT \citep{devlin2019bert} completely redefined the landscape.

In Vietnam, the release of PhoBERT \citep{nguyen2020phobert} in 2020 marked a significant turning point. Trained on a
massive 20GB corpus of Vietnamese news and Wikipedia entries, PhoBERT set new
state-of-the-art standards for tasks ranging from part-of-speech tagging to named entity
recognition. However, because PhoBERT was built on such formal text, it faces a clear
``domain gap'' when applied to social media. The messy reality of online comments, filled with
slang, teencode, and informal grammar, often proves too difficult for these traditional models
to handle.

The latest trend involves using Large Language Models (LLMs) with strong
instruction-following capabilities, such as Alibaba's Qwen series. Qwen2.5 \citep{Yang2025Qwen25} has emerged as a
particularly strong candidate because it offers exceptional support for the Vietnamese
language. Even when compared to Llama \citep{llama} or Gemma \citep{Mesnard2024Gemma} at the same parameter size, Qwen 
consistently delivers better results due to its exposure to a more diverse and high-quality
training dataset.

\subsection{Hate Speech Datasets in Vietnam}
The foundation of any AI system is its data. In Vietnam, the ViHSD (Vietnamese Hate Speech
Dataset) \citep{10.1007/978-3-030-79457-6_35} was the first large collection of data to be made public, containing over 30,000
comments labeled into three categories: Clean, Offensive, and Hate. This dataset has been
crucial for local research.

However, to better understand these interactions, the ViTHSD \citep{vithsd} (Vietnamese Targeted Hate
Speech Dataset) was created. While it builds on the data from ViHSD, it adds an important
new detail: the ``Target.'' This approach is similar to Aspect-Based Sentiment Analysis (ABSA) \citep{Pontiki2014SemEval4},
where the model doesn't just identify ``what the emotion is'', but also ``what the emotion is
directed at''.

In ViTHSD, each comment is checked to see if it targets an Individual, a Group, a Religion, an
Ethnicity, or a Political entity, and it also measures how strong the hate is. This detailed
information is important for creating smart moderation systems that can keep online spaces
safe without stopping healthy discussions.

\subsection{Explainable AI and Chain-of-Thought Prompting}
Explainability is becoming a must-have feature in modern AI. In the field of hate speech
detection, the HARE framework \citep{Yang2023HARE} introduced at EMNLP 2023 proposed a fresh approach by
using Large Language Models to generate written explanations. Instead of relying only on
classification labels, the model is trained on both the labels and the specific reasoning behind
them.

The core technique for creating these high-quality explanations is Chain-of-Thought
prompting \citep{Wei2022ChainOfThought}. Research from Google and OpenAI has shown that asking a model to think step by
step activates its internal reasoning abilities. This allows the model to solve complex problems
where standard prompting methods often fail.

In the context of the ViTHSD dataset, we use this technique to force the model to analyze
sentence structures, identify offensive terms, and pinpoint the target before it reaches a final
conclusion. This approach prevents the model from simply guessing based on keywords and
encourages a much deeper understanding of the actual meaning of the text.

\section{Analyzing the ViTHSD dataset}

\subsection{Structure and labels definition}
The ViTHSD dataset was built based on 10,000 selected and cleaned comments from the
ViHSD dataset. The labels of ViTHSD are the combination of Target and Level, represented in
the format [target]\#[level].

\textbf{Targets:}
\begin{enumerate}[leftmargin=*]
\item Individuals: Comments targeting a specific individual. This is the most common form of
online abuse.
\item Groups: Comments directed at organizations, communities, groups, or fandoms.
\item Religion: Comments targeting religious beliefs, religious figures, or religious
communities.
\item Race/Ethnicity: Comments targeting regions, ethnicities, or origins. This is the most
sensitive and dangerous category.
\item Politics: Comments related to political views, political parties, government policies, or
political figures.
\end{enumerate}

\textbf{Levels:}
\begin{itemize}[leftmargin=*]
\item Normal (0): A normal comment that does not contain negative content and does not
target any specific entity.
\item Clean (1): The comment mentions a target, but the content is clean, neutral, or positive.
\item Offensive (2): Contains vulgar, insulting, or mocking language, but does not reach the
level of hate speech or incitement to violence.
\item Hate (3): Hateful language that attacks or dehumanizes targets, incites violence,
promotes discrimination, or uses severe abusive expressions.
\end{itemize}

A key feature of ViTHSD is its multi-label nature. A comment may contain multiple targets, and
each target can have a different level. For example:
\foreignlanguage{vietnamese}{\textit{``Thằng A ngu thế cũng làm ca sĩ
(\texttt{Individuals\#Offensive}), bọn fan của nó thì cũng súc vật không kém
(\texttt{Groups\#Hate})''}} 

\subsection{Statistical Analysis of Data Imbalance}
When analyzing the dataset, we obtain the following detailed distribution table:

\begin{table}[ht]
\centering
\setlength{\tabcolsep}{3pt}
\footnotesize
\caption{Distribution of samples by Target (Train/Dev/Test)}
\label{tab:target_dist}
\resizebox{\columnwidth}{}{%
\begin{tabular}{lrrrrr}
\toprule
\textbf{Target} & \textbf{Train} & \textbf{Dev} & \textbf{Test} & \textbf{Total} & \textbf{\% } \\
\midrule
Individuals    & 5,480 & 938 & 1,398 & 7,816 & $\sim$60\% \\
Groups         & 2,977 & 517 & 769   & 4,263 & $\sim$33\% \\
Race/Ethni.    & 502   & 74  & 129   & 705   & $\sim$5\%  \\
Politics       & 363   & 57  & 89    & 509   & $\sim$4\%  \\
Religion       & 24    & 8   & 6     & 38    & $<$0.3\%   \\
\bottomrule
\end{tabular}
}
\end{table}
Our data reveals a significant imbalance across categories. The Individuals group makes up
the vast majority of the dataset, reflecting the reality of social media where personal disputes
are most frequent. In contrast, the Politics and Religion categories have very few samples.
This imbalance creates a major hurdle: traditional machine learning models tend to bias
toward the majority class. As a result, while these models may become effective at spotting
personal attacks, they often struggle with hateful comments regarding religion or politics
because they haven't seen enough examples to generalize properly. The Religion category is
particularly challenging; with only 24 training samples, it is nearly impossible for a standard
deep learning model to converge without employing specialized techniques like Data
Augmentation or Few-shot Learning.

\begin{table}[ht]
\centering
\caption{Distribution by Intensity (Level)}
\label{tab:level_dist}
% 1. Giảm khoảng cách giữa các cột (mặc định là 6pt, giảm xuống 2pt)
\setlength{\tabcolsep}{2.5pt} 
% 2. Dùng cỡ chữ nhỏ (footnotesize hoặc scriptsize)
\footnotesize 
% 3. Ép bảng theo chiều ngang của cột
\resizebox{\columnwidth}{}{%
\begin{tabular}{lrrrrr}
\toprule
\textbf{Dataset} & \textbf{Normal} & \textbf{Clean} & \textbf{Offen.} & \textbf{Hate} & \textbf{Total} \\
\midrule
Train & 1,520 & 2,480 & 1,169 & 1,831 & 7,000 \\
Dev   & 263   & 454   & 189   & 295   & 1,201 \\
Test  & 402   & 618   & 324   & 456   & 1,800 \\
\midrule
\textbf{Total} & \textbf{2,185} & \textbf{3,552} & \textbf{1,682} & \textbf{2,582} & \textbf{10,001} \\
\bottomrule
\end{tabular}
}
\end{table}
Even though the distribution by Levels is more balanced than by Targets, the CLEAN labels
still have the highest percentage. This requires our models to have a good Recall to avoid
missing the case HATE labels hidden among a large number of CLEAN comments.

\subsection{Implied Statement problem}
An analysis of the ViTHSD dataset reveals that implicit hate speech is widespread. These
comments avoid direct profanity but still carry a deeply offensive message through underlying
meaning.
\begin{itemize}[leftmargin=*]
\item Example 1: \foreignlanguage{vietnamese}{\textit{``Optimusthree với 99\% sức mạnh''}}.
\begin{itemize}[leftmargin=*]
\item Surface level analysis: it shows no offensive words, looks like a compliment about
strength.
\item Real-world context: This is a sarcastic statement in the gaming community,
implying that the player is very weak or performing poorly (only 1\% capability left).
If the model does not understand this ``meme'' context, it will label it as CLEAN.
\end{itemize}
\item Example 2: \foreignlanguage{vietnamese}{\textit{``Lũ ba que xỏ lá.''}}
\begin{itemize}[leftmargin=*]
\item Analysis: ``Ba que'' is a political slang term (referring to the yellow flag with three
red stripes of the former regime), and \foreignlanguage{vietnamese}{\textit{``xỏ lá''}} is an adjective meaning deceitful or
tricky. Combined, this forms a hate speech statement targeting a political group
(\texttt{Politics\#Hate}).
\end{itemize}
\end{itemize}

To resolve this, we added two information fields for labels marked as hate or offensive.
\begin{enumerate}[leftmargin=*]
\item Implied Statement: A short paraphrase that reveals the speaker's actual intent (For
example: \foreignlanguage{vietnamese}{\textit{``người chơi này rất tệ''}}, \foreignlanguage{vietnamese}{\textit{``nhóm người này phản động''}}).
\item Rationale: a logical explanation of why the comment is hateful based on text evidence.
\end{enumerate}
The next section explains the process for creating this information, which is the core of the
research methodology.

\section{Methodology}
We will use a pipeline that uses commercial LLMs to generate data and open-source LLMs for
training and deployment under the HARE framework.

\subsection{Designing prompts and generating reasoning data.}
Creating high-quality rationales and implied statements for thousands of comments is a
massive undertaking—doing it all by hand is simply too expensive and time-consuming to be
practical. To solve this, we used Gemini 2.0 Flash \citep{google2025gemini2flash} to act as our automated annotator. We
did not just use the model out of the box; we developed a rigorous Prompt Engineering
workflow to ensure the quality of the data. By iterating through three distinct versions of our
instructions, we were able to fine-tune the model's output until it reached the level of detail
and accuracy required for our framework.

\subsubsection{Phase 1: Prompt base}
\begin{itemize}[leftmargin=*]
\item Design: The simple structure includes a Role, a Task, and an Input/Output format. The
model is required to ``summarize the main idea'' and ``explain why.''
\item Result: The results of this phase of testing did not meet the expected requirements.
\begin{itemize}[leftmargin=*]
\item Implied Statement: The model often just paraphrases the original sentence
without revealing the hidden meaning.
\item Rationale: The explanations are too long and unfocused, not centered on the
hate-related keywords. The output format is inconsistent, making automated
processing difficult. The ability to distinguish between Hate and Offensive content
is still limited.
\end{itemize}
\end{itemize}

\subsubsection{Phase 2: Prompt updated}
\begin{itemize}[leftmargin=*]
\item Improvement: Add constraints and few-shot examples. Require the model to take the
role of a ``psychologist'' to analyze the ``hidden meaning.''
\item New issue: Domain Drift. Taking the psychologist role causes the model to analyze too
deeply from philosophical and behavioral psychology perspectives, leading to
mislabeling (false positives) for sentences that only express anger (Angry) but are not
hate speech. In addition, the model uses too many tokens for unnecessary grammatical
analysis.
\end{itemize}

\subsubsection{Phase 3: Final prompt}
This is the officially used version, integrating the HARE framework with a strict four-step
reasoning structure:
\begin{enumerate}[leftmargin=*]
\item Role \& Objective: Clearly define the task as hate speech detection.
\item Definitions: Provide precise definitions of Hate / Offensive / Clean / Normal based on
the ViTHSD standard.
\item Explicit Constraints: Enforce length limits (Implied statement: 3–8 words) and prohibit
hallucination.
\item Reasoning Framework (Logic Flow): require the model to strictly follow the flow:
Target \textrightarrow\ Implied \textrightarrow\ Evidence \textrightarrow\ Verdict.
\begin{itemize}[leftmargin=*]
\item Target: Identify the target.
\item Implied: Describe the hidden meaning or bias.
\item Evidence: Extract words or phrases with insulting or degrading meaning.
\item Verdict: Decide the label based on the evidence.
\end{itemize}
\item Output Format: Strict JSON formatting to ensure technical consistency and ease of parsing.
\end{enumerate}

\subsection{Building rationale dataset}
Data generated by the LLM cannot be guaranteed to be 100\% accurate. To remove noise, we
apply a two-step filtering process:
\begin{enumerate}[leftmargin=*]
\item Manual Verification: A random subset is reviewed by humans (5 team members) to
evaluate the quality of the Implied Statement. This helps to fine-tune the prompt one
last time.
\item Rationale Consistency Check: This is the most critical step for automated data cleaning.
\begin{itemize}[leftmargin=*]
    \item Input: Feed the Original Content + Generated Rationale back into the Gemini Flash model.
    \item Task: Verify if the rationale logically supports the ground truth label.
    \item Logic:
\begin{itemize}[leftmargin=*]
\item If Predicted Label == Ground Truth Label \textrightarrow\ Keep the
Rationale. This shows that the rationale is logical and leads to the correct
conclusion.
\item If Predicted Label $\ne$ Ground Truth Label \textrightarrow\ Discard the Rationale. This
rationale may be off-topic or contain flawed reasoning.
\end{itemize}
\item Result: From thousands of generated rationales, we filtered a high-quality
dataset (Rationale Dataset) consisting of 1,221 sample pairs for training purposes.
\end{itemize}
\end{enumerate}

\subsection{Model and Two-Stage training strategy}
We selected Qwen2.5-3B-Instruct as our backbone model. As a 3-billion-parameter,
decoder-only model, it strikes a good balance between efficiency and performance. It is
lightweight enough to be deployed on mid-range GPUs (such as the NVIDIA T4 in Kaggle or
Colab environments), yet it consistently outperforms its peers in understanding the
Vietnamese language and following complex instructions.

To ensure the model could handle both classification and reasoning effectively, we
implemented a Two-Stage Training strategy:

\textbf{Stage 1: Classifier Training}
\begin{itemize}[leftmargin=*]
\item Objective: Train the model to distinguish Target and Level classes based on the original
text.
\item Data: The full ViTHSD Train set (7,000 samples). Due to data imbalance, we applied
oversampling (+540 samples) for minority classes (Politics, Religion) to reduce model
bias.
\item Technique: Use QLoRA (Quantized Low-Rank Adaptation) \citep{dettmers2023qlora} in 4-bit. QLoRA allows
fine-tuning a large model without updating all parameters, saving GPU memory while
maintaining performance.
\item Hyperparameters: 2 epochs, Batch size 4, Learning rate 2e-4.
\end{itemize}

\textbf{Stage 2: Semantic Alignment}
\begin{itemize}[leftmargin=*]
\item Objective: Teach the model to understand the relationship between the original text
and its hidden meaning (Implied statements).
\item Data: Rationale Dataset (1,221 Content–Implied pairs).
\item Technique: Further fine-tune the model (checkpoint from Stage 1) using the rationale
data. This process adjusts the embedding space, bringing implied sentences closer to
their true meaning. For example, the model learns that in a gaming context, ``99\%
power'' $\approx$ ``weak/failed.''
\item Impact: This is the breakthrough step that allows the model to move beyond rote
keyword learning and toward deep semantic understanding.
\end{itemize}

\section{Experiments and results}

\subsection{Experimental Setup}
To evaluate the effectiveness of the proposed method, we compared it with the strongest
baseline models currently available in Vietnamese NLP and generative AI:
\begin{enumerate}[leftmargin=*]
\item PhoBERT-base \citep{nguyen2020phobert}: State-of-the-art encoder-only model for Vietnamese. This represents
the traditional fine-tuning approach.
\item Flan-T5-base \citep{chung2022scaling}: Encoder–Decoder model specialized in instruction tuning.
\item Qwen2.5-3B (Vanilla) \citep{Yang2025Qwen25}: Qwen model with standard fine-tuning only (Stage 1), without
any Rationale data.
\item Qwen2.5-3B + Rationales: Proposed model with the two-stage process.
\end{enumerate}
The main metrics used are F1-Score Micro (overall evaluation) and F1-Score Macro (fair
evaluation across classes), along with Precision and Recall.

\subsection{Quantitative Results and Comparative Analysis}
Based on the experimental results, we obtained the following metrics on the Test set:

\begin{table}[ht]
\centering
\caption{Overall performance comparison of the models (Multi-label)}
\label{tab:overall_perf_expanded}
\setlength{\tabcolsep}{2pt} % Thu hẹp khoảng cách cột tối đa
\scriptsize % Sử dụng font chữ nhỏ để chứa được nhiều dữ liệu hơn
\begin{tabularx}{\columnwidth}{lXXXX}
\toprule
\textbf{Metrics} & \textbf{Qwen (Rat)} & \textbf{Qwen (Van)} & \textbf{PhoBERT} & \textbf{Flan-T5} \\
\midrule
F1 Samples      & \textbf{0.6345} & 0.6332          & 0.5801 & 0.5074 \\
F1 Micro        & \textbf{0.6026} & 0.5900          & 0.5412 & 0.4684 \\
F1 Macro        & 0.3035          & \textbf{0.3310} & 0.2586 & 0.1311 \\
\midrule
Prec. Samples   & \textbf{0.6515} & 0.6469          & 0.5791 & 0.5033 \\
Prec. Micro     & \textbf{0.6347} & \textit{0.6100} & 0.5620 & 0.4810 \\
Prec. Macro     & \textbf{0.4067} & \textit{0.3800} & 0.3200 & 0.1800 \\
\midrule
Rec. Samples    & 0.6344          & \textbf{0.6372} & 0.5993 & 0.5210 \\
Rec. Micro      & \textbf{0.5735} & \textit{0.5600} & 0.5310 & 0.4520 \\
Rec. Macro      & \textbf{0.2842} & \textit{0.2700} & 0.2450 & 0.1100 \\
\bottomrule
\end{tabularx}
\end{table}

\textbf{Analysis:}
\begin{itemize}[leftmargin=*]
\item Superior Micro F1: The proposed model achieves 60.26\%, the highest among all. It
surpasses PhoBERT by +6.14\% and Flan-T5 by +13.42\%, confirming the advantage of a
modern decoder-only architecture combined with rationale data over older
architectures.
\item Impressive Precision: With Micro Precision of 63.47\%, the model significantly
reduces false positives. In content moderation, high precision is crucial to avoid wrongly
blocking harmless user comments, ensuring a better user experience.
\item Impact of Rationales: Compared to Qwen Vanilla, the rationale-enhanced version
improves Micro F1 (+1.26\%)and Precision (+2.47\%). While the overall increase is
modest, the real improvement lies in the hard-to-classify label classes (analyzed in
Section 5.3).
\end{itemize}

\begin{table}[ht]
\centering
\caption{Detailed F1-Score analysis by Label}
\label{tab:label_f1}
\setlength{\tabcolsep}{2pt}
\scriptsize
\begin{tabularx}{\columnwidth}{>{\raggedright\arraybackslash}p{1.8cm}XXXX} % Ép cột Label rộng 1.8cm và xuống dòng
\toprule
\textbf{Label} & \textbf{Qwen (Rat)} & \textbf{Qwen (Van)} & \textbf{Pho-BERT} & \textbf{Flan-T5} \\
\midrule
Normal & 0.7984 & 0.7970 & 0.7531 & 0.6732 \\
Indivi. \#Hate & 0.6166 & 0.6190 & 0.5365 & 0.3583 \\
Groups \#Hate & 0.5344 & 0.4583 & 0.4603 & 0.3493 \\
Politics \#Hate & 0.5400 & 0.3659 & 0.2913 & 0.0299 \\
Race \#Hate & 0.2817 & 0.3288 & 0.2857 & 0.0312 \\
\bottomrule
\end{tabularx}
\end{table}

\section{In-depth Discussion of Findings}
\begin{enumerate}[leftmargin=*]
\item Breakthrough in Detecting Political Hate (Politics):
This is the most remarkable result of the study. The Politics\#Hate label is one of the
hardest due to its high use of metaphors, political slang, and reliance on background
knowledge.
\begin{itemize}[leftmargin=*]
\item PhoBERT: 29.13\% F1
\item Qwen Vanilla: 36.59\% F1
\item Qwen + Rationales: 54.00\% F1
\end{itemize}
The F1-score improvement of +17.41\% over Qwen Vanilla and +24.87\% over PhoBERT
demonstrates that incorporating Implied Statements helped the model understand the
semantics of political comments. The model learned to associate political slang (e.g., \foreignlanguage{vietnamese}{\textit{``ba que,'' ``bò đỏ''}}) with hate meaning instead of ignoring them.

\item Improvement on Groups:
Similarly, for the Groups\#Hate label, the proposed model achieves 53.44\%, outperforming
all baselines. This demonstrates better generalization for targets that are collective groups.

\item Failure on the Race Label:
In contrast to the positive results above, performance on Race\#Hate is very low (28.17\%),
even below Qwen Vanilla.
\begin{itemize}[leftmargin=*]
\item Reason: The issue lies in the data. The Race group has only about 500 training samples.
The generated rationales for this group are also few and may contain noise. With such
data sparsity, additional fine-tuning with rationales can lead to overfitting or confuse
the model with complex reasoning that lacks enough examples for verification.
\item Key lesson: Chain-of-Thought (CoT) reasoning cannot fully replace sufficient original
data.
\end{itemize}

\item Why PhoBERT Underperforms:
PhoBERT was pre-trained on Wikipedia data (formal text). It excels at standard syntax but
lacks the flexibility to understand online language. In contrast, Qwen2.5 is a generative
model designed to understand and produce natural language, giving it a natural advantage in
capturing semantic nuances and pragmatics of modern Vietnamese.
\end{enumerate}

\section{Practical Application: Real-Time Streaming System}
To demonstrate the model's feasibility in a real-world environment, the research team
developed a demo web application that analyzes comments from YouTube. The application
focuses on simplicity and fast processing speed

\subsection{Tech Stack}
The system is developed based on a modern Client-Server model, ensuring separation
between the interface and processing logic:
\begin{itemize}[leftmargin=*]
\item Frontend (Interface): Built with React combined with Vite to optimize page load
performance. The interface is designed to be intuitive, allowing users to input a video
link and view analysis results in real-time.
\item Backend (Processing): Built on the FastAPI platform (Python). The backend acts as a
bridge, performing two main tasks: (1) Calling the YouTube Data API v3 to fetch data,
and (2) Running the trained Qwen2.5-3B model to classify comments (Inference).
\item Data Source: Integrated with YouTube Data API v3 to collect public comment data
from designated videos.
\end{itemize}

\subsection{Label Mapping Strategy}
The original labeling system of ViTHSD (\#[Level]) is academically focused and might confuse
general users. Therefore, in the demo version, we implement a mapping strategy to group the
labels into 3 simpler, more understandable categories:
\begin{itemize}[leftmargin=*]
\item Neutral: Includes comments labeled as Normal and Clean in the original system.
These are safe content that does not require moderation.
\item Offensive: Corresponds to the Offensive level. These are rude, uncomfortable
comments, but they do not necessarily violate community standards severely.
\item Hate: Corresponds to the Hate level. This is the most dangerous group, containing
inflammatory, hostile language that needs a red-flag warning system.
\end{itemize}

\textbf{Processing Workflow:}
Input URL \textrightarrow\ Crawl Comments \textrightarrow\ Qwen2.5 Inference \textrightarrow\ Map Labels \textrightarrow\ Display Result.

\subsection{Keyword Highlighting Feature (Rule-based)}
To help users quickly understand why a comment is flagged as Hate or Offensive, the demo
integrates a highlighting feature that operates on a rule-based mechanism:
\begin{itemize}[leftmargin=*]
\item Principle: The system maintains a ``blacklist dictionary'' containing hate keywords,
profanity, and commonly used slang, directly extracted from the frequency of
occurrences in the ViTHSD training dataset.
\item Functionality: When displaying comments on the React interface, the system
automatically scans and highlights words that match the dictionary entries.
\item Value: This helps moderators quickly identify ``dangerous'' areas in the text without
having to read every single word. It serves as a visual aid alongside the AI’s predicted
labels.
\end{itemize}

\section{Conclusion, Limitations, and Future Directions}
The study has demonstrated the effectiveness of combining Qwen2.5 with the HARE
framework for the ViTHSD problem. However, through the process of building the demo and
conducting experiments, we identified several important limitations that need to be
addressed:

\subsection{Limitations of the Rule-based Highlighting Mechanism}
Although the keyword highlighting feature in the demo helps with quick visualization, it
exposes inherent weaknesses of the rule-based method:
\begin{itemize}[leftmargin=*]
\item Rigidity and Lack of Context: This mechanism only detects explicit keywords. It is
entirely ineffective against implicit hate speech that doesn’t use offensive language
(e.g., \foreignlanguage{vietnamese}{\textit{``Khôn như bạn quê tôi xích đầy''}}).
\item False Positives: It can incorrectly highlight harmless words if they are in the blacklist
but are used in a neutral context, leading to unnecessary noise for the user.
\end{itemize}

\subsection{Future Directions: Attention-based Highlighting}
To address the above limitation, the next research direction will focus on developing an
Attention-based Highlighting mechanism:
\begin{itemize}[leftmargin=*]
\item Instead of using a static dictionary, the system will extract Attention Weights from the
Self-Attention layers of the Qwen2.5 model.
\item This allows the system to accurately highlight the words/phrases that the model is
actually ``focusing on'' to make its decision, including metaphorical or sarcastic
expressions, fully synchronizing the AI's reasoning with the interface display.
\end{itemize}

\subsection{Other Expansion Directions}
\begin{itemize}[leftmargin=*]
\item Addressing Data Imbalance: Collect more data or use Generative Data
Augmentation techniques for underrepresented groups, such as Race and Religion.
\item Multimodal Expansion: Extend the system to handle both images (e.g., memes) and
videos, creating a more comprehensive content moderation solution.
\end{itemize}
\bibliography{custom} 
\bibliographystyle{acl_natbib}
\end{document}
